\documentclass[10pt]{beamer} 

\usefonttheme{serif}

\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{amsmath,amssymb}
\usepackage{verbatim}

\renewcommand\rmdefault{pplx}
\renewcommand\mathfamilydefault{cmr}

\title{AWS: IAM, S3, Spot Instances and more}
\author{\underline{Xin Yin}}
\date{\scriptsize 10/01/2015\\
Cloud Computing Working Group}

\begin{document}

\frame{\titlepage}

\section{IAM - Identity and Access Management}
\begin{frame}
    \frametitle{IAM - Identity and Access Management}
\end{frame}

\section{Storage of your data}
\begin{frame}
    \frametitle{Data Storage - Your options}
    \begin{itemize}
        \item EBS - Elastic Block Store 
            \begin{itemize}
                \item {\color{green} Good:} can mounted as a block device to EC2 instances.
                \item {\color{green} Good:} provides file system.
                \item {\color{red} Bad:} can only be mounted to one instance.
                \item {\color{red} Bad:} charged at provisioned volume.
            \end{itemize}
        \item S3 - Simple Storage Service 
            \begin{itemize}
                \item {\color{green} Good:} {\it Simple} and versatile
                \item {\color{green} Good:} more reliable (than EBS)
                \item {\color{green} Good:} cheaper (than EBS)\\ {\it $\$ 0.03$/GiB~ vs. $~\$ 0.10$/GiB (provisioned)}
                \item {\color{green} Good:} supports versioning
                \item {\color{green} Good:} files can be made publicly available 
                \item {\color{red} Bad:} not a file system 
            \end{itemize}
    \end{itemize}
\end{frame}
\subsection{EBS - Elastic Block Store}

\begin{frame}
    \frametitle{S3 - Simple Storage Service}
    \begin{itemize}
        \item live demo
    \end{itemize}
\end{frame}

\section{EC2 (Spot) Instances}
\begin{frame}
    \frametitle{Calculate the costs of your spot instances}
    \begin{itemize}
        \item EC2 Instances: price varies by instance type
        \item S3 Storage: 
            \begin{itemize}
                \item Volume: {\color{blue} $\$$ 0.03 / GiB $\times$ month}
                \item Requests: {\color{blue} $\sim$ $\$$ 0.005 / 1000 Reqs}
            \end{itemize}
        \item Data transfer: 
            \begin{itemize}
                \item Inbound: {\it free}
                \item Outbound (EC2/S3 to the Internet): {\color{blue} \it $\$$ 0.09 / GiB}
                \item Between AWS Regions: {\color{blue} \it $\$$ 0.02 / GiB}
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{How to reduce costs?}
    \begin{enumerate}
        \item Determine the most suitable instance types
        \item Minimize/eliminate cross-regional data transmission
        \item Maximize the utilization of computational resources
        \item Protect yourself from data loss due to instance termination
    \end{enumerate}
\end{frame}


\begin{frame}[fragile]
    \frametitle{What is the best instance type for me?}
    \begin{itemize}
        \item Determing the limiting factor of your program:
            \begin{itemize}
                \item CPU: {\it simulations}
                \item Memory: {\it bioinformatics}
                \item Storage: {\it big datasets}
            \end{itemize}
        \item {\tt time}
            {\scriptsize
                \begin{verbatim}
alias timem="/usr/bin/time -f 'real \%e\nuser \%U\nsys \%S\nmem \%M Kb'"
                \end{verbatim}
            }

            \begin{verbatim}
    timem <some_program>
    real 18390.10
    user 128669.82
    sys 2318.05
    mem 45031068 Kbytes 
            \end{verbatim}
            \begin{itemize}
                \item $(128669.82 + 2318.05) / 18390.10 = 7.13  \approx\text{(8 cores)}$
                \item $45031068 / 1024 / 1024 = 42.94 \text{(GiB)}$
            \end{itemize}
        \pause
        \item How many GBs of memory per core?
        \item \url{https://aws.amazon.com/ec2/instance-types/}

    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Minimize cross-regional data transmission}
    \begin{itemize}
        \item Launch EC2 instances in the same region of your S3 buckets
            \begin{itemize}
                \item No {\it \$ 0.02 / GiB} cost for data transfer
                \item Low-latency, high-throughput
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Maximize the utilization of computational resources}
    \begin{itemize}
        \item CPUs, are they 100\% utilized? 
        \item Reduce the amount of time on setups:
            \begin{itemize}
                \item Use {\it user data}
                \item Launch instance from your AMI
                \item Use {\bf CodeDeploy}
            \end{itemize}
        \item Load balancing: 
        \item Automation and monitoring:
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Inevitable price fluctutation}
    \begin{itemize}
        \item Study the pricing history (and predict the variation)
        \item Don't be intimidated to bid {\bf higher}
            \begin{itemize}
                \item You are only charged for the {\it current price}, not your {\it maximum bid price}.
                \item Protect you from price surges and spikes.
            \end{itemize}
        \item If possible, divide your program into small segments:
            \begin{itemize}
                \item Example: simulations, MCMC
                \item After each segment, backup the results to S3
            \end{itemize}
        \item Termination Notices:
            \\
            \url{http://169.254.169.254/latest/meta-data/spot/termination-time}
    \end{itemize}
\end{frame}

\end{document}
